### 2020-05-04 tour-vis.Rmd
init a distill article

---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: Sherry Zhang 
    url: https://github.com/huizezhang-sherry
    affiliation: Monash University
    affiliation_url: https://www.monash.edu/
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.