### 2020-05-04 tour-vis.Rmd
lit review and global structure

---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: Sherry Zhang 
    url: https://github.com/huizezhang-sherry
    affiliation: Monash University
    affiliation_url: https://www.monash.edu/
date: "`r Sys.Date()`"
output: distill::distill_article
bibliography: biblio.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
load(here::here("data/temp.rda"))
```

# Introduction

## Tour

Tour provides a way to explore multivariate data interactively via an established tour path. A tour path is formed by interpolating between randomly generated plane. Different types of tours are available depends on the purpose of the exploration i.e. a grand tour is suitable for randomly exploring the data from different angles; a guided tour detects a particular structure in the data. a manual tour allows user to manually control the projection @cook1997manual.

## Guided tour

Guided tour is usually used in conjunction projection pursuit, a method coined by @friedman1974projection to detect "interesting" low-diemsion projection of multivariate data. A projection pursuit requires the definition of a projection pursuit index function and an optimisation routine. The projection pursuit index measures the "interestingness" of data defined as its departure from normality. Numerous indices have been proposed in the literature, including lengendre index @friedman1974projection, hermite index @hall1989polynomial, natural hermite index @cook1993projection, chi-square index @posse1995projection, LDA index @lee2005projection and PDA index @lee2010projection. An optimisation routine is required to find the projection basis (thus projection) that maximises the projection pursuit index. The discussion of existing optimisation procedure will be discussed in the next section.

Guided tour creates visualisation for the projections found by projection pursuit by constructing a tour path. An illustration from @buja2005computational entails the geodesic interpolation bewteen the plane generated by projection pursuit.

![](figures/tour-path.png)

## Optimisation methods existing in tourr

Because of the fact that projection pursuit index can be non-differentiable and that the projection basis is constrained to be orthonormal, the optimisation need to be a derivative-free, constrained procedure. Also, local maxima is equally interested as global maxima since it reveals distinct local structure in the data. Simulated annealing, proposed by [@posse1995projection], is widely used to optimise the projection pursuit index. Geodesic searching method is an analogue to gradient acsent with the evaluation of derivative being replaced by randomly evaluation five direction in a tiny nearby neighbourhood and then find the projection basis along the geodesic direction that maximise the projection pursuit index. [need refine and more details here!]

## problems experienced with existing versions

# Recording the guided tour

## Tour components

Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) @wickham2011tourrpackage. A tour includes two major components: a *generator* that generating the projection basis according to projection pursuit and an *interpolator* that performing geodesic interpolation between the projection basis.

The psudo-code below illustrates the implementation of guided tour in the tourr package. Given an projection pursuit index function and a randomly generated projection basis (current basis), the optimisation procedure produces a target basis inside `generator()`  Both the current basis and the target basis will be supplied to `tour_path()` to prepare information needed for constructing a geodesic path. This information is then used to compute a series of interpolating bases inside the `tour()` function. All the basis will be sent to create animation for visualising the tour in the `animate()` function.

```{r eval = FALSE, echo = TRUE}
animation <- function(){

  # compute projection basis 
  tour <- function(){

    # construct bases on the tour path
    new_geodesic_path <- function(){
      tour_path <- function(){

        # GENERATOR: generate projection basis via projection pursuit
        guided_tour <- function(){
          generator <- function(){

            # define projection pursuit index
            # generate the target basis from the current basis via optimisation
          }
        }

        # prepare geodesic information needed for interpolating along the tour path
      }
    }

    # INTERPOLATOR: interpolate between the current and target basis 
    function(){
      # generate interpolating bases on the geodesic path
    }
  }

  # animate according to different display methods
}
```

## Data Structure

In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randonly nearby bases generated in the optimisation are not stored. This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario [needa rephrase this part].

The global structure consists of six columnes: `basis`, `index_val`, `tries`, `info`, `loop`, `id` and captured all the basis generated during whole tour process. The first five rows of the record of a five-variable 1d projection is printed below.

```{r eval = FALSE}
devtools::load_all()
set.seed(123456)
temp <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
                                           search_f =  search_geodesic),
                  rescale = FALSE, verbose = TRUE)

#save(temp, file = here::here("vignettes", "data", "temp.rda"))
```

```{r echo = TRUE}
temp %>% head(5)
```

`tries` has an increment of one once the generator is called (equivalently a new target basis is generated); `info` records the stage the basis is in. This would include the `interpolation` stage and the detailed stage in the optimisation i.e. `direction_search`, `best_direction_search`, `line_search`and `best_line_search` for geodesic searching (`search_geodesic`); `random_search` and `new_basis` for simulating annealing (`search_better`). `loop` is the counter used for the optimisation procedure and thus will be `NA` for interpolation steps. `id` creates a sequential order of the basis. This information will be stored and printed when the optimisation ends and can be turned off via `print = FALSE`. Additional messages during the optimisation can be displayed via `verbose = TRUE`.

# Visual methods
## index vs time plots
## target basis vs interpolation steps, index vs time
## low-d representation, PCA

# Animating plots

# Finding errors and developing improvements
## interrupt
## polish

# Vis package