### 2020-05-13 source/cheatsheet.Rmd
supplementary materials

---
title: "Performance of different optimisation routines on projection pursuit index"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ferrn)
library(patchwork)

load(here::here("source/test/holes_1d_geo.rda"))
load(here::here("source/test/holes_1d_better.rda"))
load(here::here("source/test/holes_2d_geo.rda"))
load(here::here("source/test/holes_2d_geo_polish.rda"))
load(here::here("source/test/holes_2d_better.rda"))
load(here::here("source/test/holes_2d_better_polish.rda"))
load(here::here("source/test/holes_2d_better_max_tries.rda"))
load(here::here("source/test/holes_2d_better_max_tries_polish.rda"))
load(here::here("source/test/holes_2d_better_random.rda"))
load(here::here("source/test/holes_2d_better_random_polish.rda"))
load(here::here("source/test/holes_2d_pos.rda"))
load(here::here("source/test/holes_2d_pos_polish.rda"))

```

# Holes Index

## 1D projection

```{r}
p1 <- holes_1d_geo %>% explore_trace_interp()
p2 <- holes_1d_better %>% explore_trace_interp()
(p1 | p2 ) & ylim(0.74, 0.95) 
```

## 2D projection

```{r}
p1 <- bind_rows(holes_2d_geo, holes_2d_geo_polish) %>% 
  explore_trace_interp(point = TRUE) + ggtitle("geodesic")
p2 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>% 
  explore_trace_interp(point = TRUE) + ggtitle("better, max.tries = 25")
p3 <- bind_rows(holes_2d_better_max_tries, holes_2d_better_max_tries_polish) %>%
  explore_trace_interp(point = TRUE) + ggtitle("better, max.tries = 500")
p4 <- bind_rows(holes_2d_better_random, holes_2d_better_random_polish) %>% 
  explore_trace_interp(point = TRUE) + ggtitle("better_random")
p5 <- bind_rows(holes_2d_pos, holes_2d_pos_polish) %>% 
  mutate(method = fct_relevel(method, c("search_posse", "search_polish"))) %>% 
  explore_trace_interp(point = TRUE) + ggtitle("posse")

((p1 | p4 | p5 ) / (p2 | p3 )) & 
  theme(legend.position = "none") & 
  ylim(0.8, 0.99)

```