### 2020-06-09 source/geozoo.Rmd
a tiny example of geozoo

---
title: "geozoo"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(geozoo)

files <- paste0("data/", list.files(here::here("data")))
purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))

```

```{r}
pca <- holes_1d_geo %>% ferrn::compute_pca()

interp <- pca$combined %>% filter(info == "interpolation") %>% dplyr::select(V1:V5) %>% as.matrix()

a <- sphere.hollow(p = 5, n = 1000)
a$points <- a$points %>% rbind(interp)
color <- c(rep("black", 1000), rep("red", nrow(interp)))
print(a, col = color)
```