### 2020-07-07 source/geozoo.Rmd
add lines between points in geozoo animation

---
title: "geozoo"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(geozoo)
library(ferrn)

files <- paste0("data/", list.files(here::here("data")))
purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))

```

```{r echo = FALSE}
vis <- function(data, color = NULL, edges = NULL, edges.col = NULL, ...){

  row <- data$basis[[1]] %>% nrow()
  col <- data$basis[[1]] %>% ncol()
  dim <- row * col
  n_bg <- 1000

  path <- data %>% pull(basis) %>% purrr::flatten_dbl() %>% matrix(ncol = dim, byrow = TRUE)

  a <- sphere.hollow(p = dim, n = n_bg)
  a$points <- a$points %>% rbind(path)

  if(is.null(color)){
    color <- c(rep("grey", 1000), rep("red", nrow(data)))
  }

  # if(is.null(edges)){
  #   start <- n_bg + 1
  #   end <- nrow(a$points)
  #   edges <- matrix(c(start: (end -1), 
  #                     (start + 1): end), ncol = 2)
  # }
  # 
  # if(is.null(edges.col)){
  #   edges.col <- "red"
  # }

  print(a, col = color, edges = edges, edges.col = edges.col)

}
```

1. 1D projection using search_geodesic

- The final projection is colored green, with rotation, it is always posed on the opposite direction of V2 with same magnitude.

```{r}
# final basis 
holes_1d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
# final basis after polish
holes_1d_geo_polish %>% tail(1) %>% pull(basis) %>% .[[1]]

# searching path and final basis
dt <- holes_1d_geo %>% filter(info == "interpolation")
color <- hue_pal()(1)
color_vec <-  c(rep("grey", 1000), rep(color, nrow(dt)))
vis(dt, color = color_vec, edges.col = color)

# searching path, final basis and polish
interp <- holes_1d_geo %>% filter(info == "interpolation")
polish <- holes_1d_geo_polish %>% filter(info == "loop_best")
dt_bind <- bind_rows(interp,polish) 
color1 <-  hue_pal()(2)[1]
color2 <-  hue_pal()(2)[2]
color_vec <- c(rep("grey", 1000), rep(color1, nrow(interp)), rep(color2, nrow(polish)))
edges.col <- c(rep(color1, nrow(interp)), rep(color2, nrow(polish)))
vis(dt_bind, color = color_vec, edges.col = edges.col)

```

2. Comparing two search better paths

```{r}
# different seeds lead to different variables found
interp1 <- kol_cdf_1d_2var_better %>% filter(info == "interpolation")
interp2 <- kol_cdf_1d_2var_better_2 %>% filter(info == "interpolation")
dt_bind <- bind_rows(interp1, interp2)

# a small example 
interp1 %>% tail(1) %>% pull(basis) %>% .[[1]]
interp2 %>% tail(1) %>% pull(basis) %>% .[[1]]

color_hex <- RColorBrewer::brewer.pal(4, "Dark2")
color_vec<- c(rep("grey", 1000), 
           rep(color_hex[[1]], nrow(interp1)),
           rep(color_hex[[2]], nrow(interp2)))
edges <- matrix(
  c((1000 + 1): (1000 + nrow(interp1)-1), (1000 + nrow(interp1) + 1): (1000 + nrow(dt_bind)-1), 
    (1000 + 2): (1000 + nrow(interp1)), (1000 + nrow(interp1) + 2): (1000 + nrow(dt_bind))), 
  ncol = 2
)
edges.col <- c(rep(color_hex[[1]], nrow(interp1)), 
               rep(color_hex[[2]], nrow(interp2)))

vis(dt_bind, color_vec, edges = edges, edges.col = edges.col)
```

```{r}
# don't think adding the lines is particularly useful tho
# explore the full parameter space 
interp3 <- kol_cdf_1d_2var_better_random %>% filter(info == "interpolation")
polish3 <- kol_cdf_1d_2var_better_random_polish %>% filter(info == "polish_best")
n_interp3 <- nrow(interp3) + nrow(polish3)

dt_bind <- bind_rows(interp3, polish3)

color_hex <- RColorBrewer::brewer.pal(4, "Dark2")
color_vec <- c(rep("grey", 1000), 
           rep(color_hex[[3]], n_interp3))

edges <- matrix(
  c((1000 + 1): (1000 + nrow(dt_bind)-1),

    (1000 + 2): (1000 + nrow(dt_bind))),

  ncol = 2
)

edges.col <- c(rep(color_hex[[3]], n_interp3))

vis(dt_bind, color_vec, edges = edges, edges.col = edges.col)
```

```{r}
# combined all three together
# better random finds global 
# explore the full parameter space 
interp1 <- kol_cdf_1d_2var_better %>% filter(info == "interpolation")
interp2 <- kol_cdf_1d_2var_better_2 %>% filter(info == "interpolation")
interp3 <- kol_cdf_1d_2var_better_random %>% filter(info == "interpolation")

polish1 <- kol_cdf_1d_2var_better_polish %>% filter(info == "polish_best")
polish2 <- kol_cdf_1d_2var_better_2_polish %>% filter(info == "polish_best")
polish3 <- kol_cdf_1d_2var_better_random_polish %>% filter(info == "polish_best")

n_interp1 <- nrow(interp1) + nrow(polish1)
n_interp2 <- nrow(interp2) + nrow(polish2)
n_interp3 <- nrow(interp3) + nrow(polish3)

dt_bind <- bind_rows(interp1, polish1, 
                     interp2, polish2,
                     interp3, polish3)

color_hex <- RColorBrewer::brewer.pal(4, "Dark2")
color_vec <- c(rep("grey", 1000), 
           rep(color_hex[[1]], n_interp1), 
           rep(color_hex[[2]], n_interp2), 
           rep(color_hex[[3]], n_interp3))

edges <- matrix(
  c((1000 + 1): (1000 + n_interp1), 
    (1000 + n_interp1 + 1): (1000 + n_interp1 + n_interp2-1), 
    (1000 + n_interp1 + n_interp2 + 1): (1000 + nrow(dt_bind)-1),

    (1000 + 2): (1000 + n_interp1 + 1), 
    (1000 + n_interp1 + 2): (1000 + n_interp1 + n_interp2), 
    (1000 + n_interp1 + n_interp2 + 2): (1000 + nrow(dt_bind))),

  ncol = 2
)

edges.col <- c(rep(color_hex[[1]], n_interp1), 
               rep(color_hex[[2]], n_interp2),
               rep(color_hex[[3]], n_interp3))

vis(dt_bind, color_vec, edges = edges, edges.col = edges.col)

# explore the reduced 2D space via PCAs
interp1 <- kol_cdf_1d_2var_better 
interp2 <- kol_cdf_1d_2var_better_2
interp3 <- kol_cdf_1d_2var_better_random

pca <- bind_rows(interp1, interp2, interp3) %>% 
  mutate(algorithm = c(rep("better1", nrow(interp1)), 
                       rep("better2", nrow(interp2)), 
                       rep("better_random", nrow(interp3))),
         algorithm = fct_relevel(algorithm, c("better1", "better2", "better_random"))) %>% 
  ferrn::compute_pca()

pca$combined %>% 
  ggplot(aes(x = PC1, y = PC2, color = algorithm)) + 
  geom_point() + 
  scale_color_brewer(palette = "Dark2") + 
  theme(aspect.ratio = 1)

interp3 %>% filter(tries > 40) %>% explore_trace_search()
interp3 %>% explore_trace_search()
```

```{r}
# finds V8
holes_2d_better %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis)
# finds V3 and V8
holes_2d_better_max_tries %>% filter(info == "interpolation")%>% tail(1) %>% pull(basis)

interp1 <- holes_2d_better %>% filter(info == "interpolation")
polish1 <- holes_2d_better_polish %>% filter(info == "polish_best")
interp2 <- holes_2d_better_max_tries %>% filter(info == "interpolation")
polish2 <- holes_2d_better_max_tries_polish %>% filter(info == "polish_best")
dt_bind <- bind_rows(interp1, 
                     polish1, 
                     interp2, 
                     polish2
                     )

color_hex <- scales::brewer_pal(palette = "Dark2")(8)
scales::show_col(color_hex)
# color <- c(rep("grey", 1000), 
#            rep(color_hex[1], nrow(interp1)-1), color_hex[2], rep(color_hex[3], nrow(polish1)-1), color_hex[4],
#            rep(color_hex[5], nrow(interp2)-1), color_hex[6], rep(color_hex[7], nrow(polish2)-1), color_hex[8]
#            )

color <- c(rep("grey", 1000), 
           rep(color_hex[1], nrow(interp1)-1), "green", rep("red", nrow(polish1)-1), "blue",
           rep(color_hex[5], nrow(interp2)-1), "green", rep("lightblue", nrow(polish2)-1), "blue"
           )

dt_bind <- bind_rows(interp2, polish2)
color <- c(rep("grey", 1000), 
           rep(color_hex[1], nrow(interp2)), rep("red", nrow(polish2)-1), "blue")
vis(dt_bind, color)

```

```{r}
pca <- holes_1d_geo %>% ferrn::compute_pca()

interp <- pca$combined %>% dplyr::select(V1:V5) %>% as.matrix()

a <- sphere.hollow(p = 5, n = 1000)
a$points <- a$points %>% rbind(interp)
color <- c(rep("black", 1000), "green",rep("red", nrow(interp)-1))
print(a, col = color)
```

$\mathcal{S}_{p-1}$ minus one because