### 2020-08-24 numbat_talk/3min_talk_slide.Rmd
numbat 3 mins talk materials

---
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      navigation:
          ratio: '16:9'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# css: ["presdefault.css"] 
```

```{r}
files <- paste0("data/", list.files(here::here("data")))
purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))

source(here::here("source/indice.r"))
source(here::here("source/sim_data.r"))

library(tidyverse)
library(ferrn)
library(patchwork)
library(ggrepel)
```

#### Diagnostics of optimisation algorithms in projection pursuit guided tour

.pull-left[

**Tour:** exploratory data analysis tool that can detect the structure of data

Data: $X_{n \times p}$

Projection: $P_{n \times d} = X \cdot A_{p \times d}$

Define a projection pursuit index: $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$

Optimisation problem:

\begin{align}
&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
&s.t.  \mathbf{A}^{\prime} \mathbf{A} = I_d
\end{align}

Three existing algorithms:

- search_better
- search_better_random
- search_geodesic

]

.pull-right[
```{r fig.height=5}
colour <- RColorBrewer::brewer.pal(3, "Dark2")
max_1 <- kol_cdf_1d_2var_better %>% filter(index_val == max(index_val)) %>% pull(index_val)
max_2 <- kol_cdf_1d_2var_better_2 %>% filter(index_val == max(index_val)) %>% pull(index_val)

p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp(col = 0) + 
  scale_colour_manual(values = colour[1]) + ggtitle("search better, seed 123456")
p6 <- kol_cdf_1d_2var_better_2 %>% explore_trace_interp(col = 0) + 
  scale_colour_manual(values = colour[2]) + ggtitle("search better, seed 12345")

p7 <- kol_cdf_1d_2var_better_random %>% 
  explore_trace_interp(col = 0) + 
  ylim(0, 0.25) + 
  scale_colour_manual(values = colour[3]) + 
  theme(legend.position = "none") + 
  ggtitle("search better random")

((p5 | p6) / p7) &
  theme(legend.position = "none") & ylim(0, 0.25) & ylab("index value")

```

```{r fig.height=5}
interp1 <- kol_cdf_1d_2var_better 
interp2 <- kol_cdf_1d_2var_better_2
interp3 <- kol_cdf_1d_2var_better_random

dt <- bind_rows(interp1, interp2, interp3) %>% 
  mutate(algorithm = c(rep("search better 1", nrow(interp1)), 
                       rep("search better 2", nrow(interp2)), 
                       rep("search better random", nrow(interp3))),
         algorithm = fct_relevel(algorithm, c("search better 1", "search better 2", "search better random"))) %>% 
  compute_pca() %>% 
  `$`(combined)

dt %>% filter(info == "interpolation") %>% group_by(algorithm) %>% filter(id == max(id)) -> b

dt %>% 
  ggplot(aes(x = PC1, y = PC2, color = algorithm)) + 
  geom_point() +
  geom_path(alpha = 0.1) + 
  geom_point(data = b, size =10) + 
  theme(aspect.ratio = 1) + 
  scale_colour_brewer(palette = "Dark2") 
  #geom_point(data = dt %>% filter(info == "start"), color = "blue") +

```

```{r eval = FALSE}
a <- interp1 %>% compute_pca() %>% `$`(combined)

a %>% ggplot(aes(x = PC1, y = PC2, color = info)) + 
  geom_point() +  
  geom_path() + 
  geom_point(data = a %>% filter(info == "start")) + 
  theme(aspect.ratio = 1)

```

]