### 2021-09-05 review/response.Rmd
finish response to reviewer

---
title: "Response"
author: "H. Sherry Zhang"
date: "22/08/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reviewer 1

## The Review

This work presents visual diagnostic tools for different methods of projection pursuit optimization. They use line plots, PCA and the tour method to visually compare the behavior of different optimizers. Finally, they showcased the analysis of three optimizers using their tool.

The tool is likely to be of interest to the R community.  Most of my comments are about the technical aspect of the content. The writing is overall good, but please see my comment under the "writing" regarding to the clarity of the text.

### Major comment

- "Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign...However, this creates difficulties for comparing the optimisers since the end bases will be symmetric to the origin. A sign flipping step is conducted..." and the "Reconciling the orientation" section
  - For 1D projections, flipping the sign is sufficient. For p-D projections of $\mathbb{R}^n$​, one may need a little more. To my knowledge most index functions are invariant to any rotations that happens within the p-subspace. This makes me wonder that in general, one may want to align two bases so that their difference does not contain any "within-subspace" components. In other words, one may want to compare representations in the Grassmannian manifold  (space of k-subspaces) instead of in the Stiefel manifold (space of orthonormal k-frames).

\textcolor{red}{With p-D bases, the correction of orientation is a bit more complicated. Take 2D as an example, two bases can be aligned in the first axis but not in the second. This would means there are $2^p$ possibility to check.}

### Minor comment

- The software seems to be specific to projection pursuit. I wish the authors could comment more on its potential generality to other optimizations. For example, can one generalizes the tour method to visualize any parameter trace of an optimizer?

\textcolor{red}{It is feasible to generalise the tour method for diagnosing other optimisers, providing the parameter space can be simulated. In our examples, the parameter to trace (projection base ($p \times d$) ) lives in an p-sphere, which can be simulated by the `geozoo` package. For other optimisers, reserachers need first to create a data collection object in a tibble that track the parameter during the optimisation, and then to use ferrn to produce the diagnostic plots as shown in the paper.}

- Figure 4 "...All the bases in PD have been flipped for easier comparison of the final bases..."
  - This is related to the major comment. In the figure, PD flipped in sign but stopped at a basis representation similar to the destination of the CRS. Flipping the sign back, does it mean that PD and CRS diverged but stopped at the similar subspace with opposite basis representations? Also, given that PD and CRS have the same starting bases and one is flipped, why wouldn't the dashed line pass the origin (gray dot) in the figure?

\textcolor{red}{Yes the reviewer is giving a decent explanation in describing the effect of flipping. The gray dot should land on the dashed line and this is a mistake due to the data not being updated after flipping in the code. This has now been corrected.}

- "Recall that the columns in a 2D basis are orthogonal to each other, so the space of *p* × 2 bases is a *p*-D torus".

  - It is not clear to me why "so". For p=2, the space of 2x2 bases is a pair of 1-D tori instead of just one (because you have two orientations for a pair of orthogonal vectors), right? For p=3, the space of 3x2 bases is SO(3), which is not a 3-D torus? Ref: https://en.wikipedia.org/wiki/Stiefel_manifold

\textcolor{red}{We change the wording of the sentence to "**a torus in the $p$-D space**"}

### Writing

- "Along with the bases recorded during the optimisation and a zero basis, the parameters (centre and radius) of the 2D space can be estimated by PCA. The centre of the circle is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."
  - It took me multiple passes to get what this means. Perhaps instead starting with something like: "We use PCA to project and visualize the parameters/bases in 2D. The centre of the 2D view is..."?

\textcolor{red}{Changed to "We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis. The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."}

- "Figure 6 shows some frames from the tour plot of the same two optimisations in its original 5D space."
  - I think this is the first mention of "5D space"/5D data in the main text (apart from the caption in Figure 4), so it may be worth clarifying what the original 5D space is.

\textcolor{red}{The explanation of data set is introduced in the subsection *Simulation setup* under *Diagnosing an optimiser*. To avoid ambiguity, we change the "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".}

- "While the previous few examples have looked at the space of 1D bases ...  2D bases"
  - I don't think the terms "1D bases" and "2D bases" are standard. Perhaps (orthonormal) "1-frames" and "2-frames"?

\textcolor{red}{Changed to "While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames."}

# Reviewer 2

These experiments are interesting and so is the methodology.
The conclusions could do with beefing up
What can we learn about how to optimise the holes index for 1 and 2d?
Does this apply to other smooth indices? What about non-smooth indices?
What about indices in other dimensions?

\textcolor{red}{A paragraph summarising the finding is added as the second paragram in the conclusion}

1. Should refer to earlier work on scagnostic indices, page 3

\textcolor{red}{Added reference of (Wilkinson et al., 2005) and  (Wilkinson and Wills, 2008):}

\textcolor{red}{L. Wilkinson and G. Wills. Scagnostics distributions. Journal of Computational and Graphical Statistics, 17 (2):473–491, 2008}

\textcolor{red}{L. Wilkinson, A. Anand, and R. Grossman. Graph-theoretic scagnostics. In IEEE Symposium on Information Visualization, 2005. INFOVIS 2005., pages 157–164, Oct 2005.}

2. figure 2 caption "using the holes index"  "and a boxplot". What is the colour scale for the points? The text related to figure 2 in the bottom of page 5 should mention the dataset and the index, giving references if needed.

\textcolor{red}{Fixed language. Added mention the color scale in the figure 2 caption:  "The color scale is from the customised botanical palette in the ferrn package."}

3. "with higher index value"  -> "with a higher index value". The word "the" is missing in a few other places in the paper.

\textcolor{red}{corrected.}

4. In Figure 3 would be better if the y-axes scales were the same for the two graphs.

\textcolor{red}{fixed}

5. What is the star in figures 4 and 5? This is not mentioned in the text or caption

\textcolor{red}{Added "The black star represents the theoretical best basis the optimisers aiming to find." in the caption of figure 4.}

6. In the text you say things like "details=TRUE" and "animate=TRUE" but what functions are you referring to?

\textcolor{red}{Mention of "details=TRUE" and "animate=TRUE" is removed from subsection diagnostic 3a and 3b to the implementation section: `explore\_space\_pca()` produces the PCA plot of projection bases on the reduced space. Figure 4 includes the additional details of anchor and search bases, which can be turned on by the argument `details = TRUE`. The animated version in Figure 5 is produced with argument `animate = TRUE`.}

7. In Figure 9, it seems CRS is doing better here than in Figure 8-- maybe this could be mentioned.

\textcolor{red}{Figure 8 and 9 focuses on different aspects: Figure 8 compares the effect of implementing an interruption while Figure 9 compares the effect of polishing. Cross comparing the two experiments is less meaningful.}

8. "not the case for all the indexes" missing a "." page 11

\textcolor{red}{added}

9. For the Kolmogorov index, why are you using randomly generated data instead of normal quantiles?

\textcolor{red}{The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on data. I am not sure what would be the reasoning for using normal quantiles in this index. That would imply you are assuming normality on the theoretial pdf.}

10. Fix references with "et al"

\textcolor{red}{fixed}